Estimating duration in the presence of misclassification

### 2024-02-20 transmission.tex
fix links

In this chapter, I use the estimate of the duration of RT-PCR positivity from combining information from the ATACCC and CIS studies (see \cref{E-imperf-test:sec:conclusion}) to investigate the feasibility of estimating incidence and transmission of SARS-CoV-2 from only the CIS data (described in \cref{E-biology-data:sec:positivity}).
I will show that the CIS data is sufficient; demonstrating the value of surveys of this type for pandemic surveillance.

@@

I keep the same logistic regression model in terms of prevalence, transforming its posterior distribution of prevalence to a posterior distribution of incidence.
This results in a two-step process: first, the same inference process described in \cref{E-biology-data:sec:MRP} is used to estimate prevalence; second, the prevalence is transformed to incidence.
An alternative would be an inference procedure that estimates incidence proportion, fit directly to the RT-PCR test results.
This is required to fulfil several desirable features of the model.
Denote the quantity to be inferred (which could be either incidence proportion or prevalence) at time $t$ in strata $i$ as $\pi_{it}$.
The strata will represent at least age group and region, although including ethnicity and sex is also due to non-response in the survey (see \cref{E-biology-data:sec:positivity}).
For many combinations of stratum $i$ and day $t$, the number of positives will be small, often fewer than 5; therefore some form of borrowing between time and strata is required.
The statistical model and inference procedure should have the following properties.
\begin{enumerate}
    \item The smoothing model over time should penalise changes from continuous exponential growth.
        Penalising changes from continuous exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalisation has previously been shown to improve estimates of SARS-CoV-2 prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}.
    \item Limit estimates to epidemiologically and biologically plausible values.
        That is, incidence proportion and prevalence should be in the interval $[0, 1]$.
        % While incidence proportion can theoretically be greater than 1, this would imply that some individuals were infected multiple times in a day which is not biologically plausible.
    \item Have sufficient parameters to incorporate all necessary covariates to correct for non-response bias (see \cref{E-biology-data:sec:positivity}).
    \item Borrow information across the different covariates.
        The borrowing should penalise towards the distribution of incidence/prevalence across strata in the same geographic region being constant.
        Standard stratified, mechanistic transmission model (such as the one introduced in \cref{E-SEIR:sec:structured-populations}) predict this, if behavioural changes are independent of strata and there is negligible impact of immunity.
    \item Be computationally efficient.
        The model will be fit to granular data and therefore needs to be run within the SRS, with its computational limitations (see\todo{ref section on SRS computational limitations}).
\end{enumerate}

@@

I link the modelled proportion of individuals that are RT-PCR positive to the data using a beta-binomial likelihood (as discussed in \cref{E-biology-data:sec:positivity}).
If on each day $t$, each strata $i$ is observed to have $y_{at}$ positives out of $n_{at}$ tests, then the likelihood is:
\begin{align}
    \prod_{a,t} \BB (y_{at} \mid n_{at}, \bar{p}_{at}, \rho)
    \label{SEIR:eq:likelihood}
\end{align}
where: $\bar{p}_{at}$ is the mean proportion of individuals in strata $a$ that are RT-PCR positive on day $t$; and $\rho$ is a nuisance parameter controlling the overdispersion of the beta-binomial distribution.
Denoting the proportion of stratum $a$ in state $p_k$ as $p_{ka}(\tau)$, then $\bar{p}_{at} = \int_{t}^{t+1} \sum_{k=1}^l p_{ka}(\tau) d\tau$.

>>>>

In this chapter, I use the estimate of the duration of RT-PCR positivity from combining information from the ATACCC and CIS studies (see \cref{E-imperf-test:sec:conclusion}) to investigate the feasibility of estimating incidence and transmission of SARS-CoV-2 from only the CIS data (described in \cref{E-intro:sec:cis}).
I will show that the CIS data is sufficient; demonstrating the value of surveys of this type for pandemic surveillance.

@@

I keep the same logistic regression model in terms of prevalence, transforming its posterior distribution of prevalence to a posterior distribution of incidence.
This results in a two-step process: first, the same inference process described in \cref{E-biology-data:sec:MRP} is used to estimate prevalence; second, the prevalence is transformed to incidence.
An alternative would be an inference procedure that estimates incidence proportion, fit directly to the RT-PCR test results.
This is required to fulfil several desirable features of the model.
Denote the quantity to be inferred (which could be either incidence proportion or prevalence) at time $t$ in strata $i$ as $\pi_{it}$.
The strata will represent at least age group and region, although including ethnicity and sex is also due to non-response in the survey (see \cref{E-intro:sec:cis-strata}).
For many combinations of stratum $i$ and day $t$, the number of positives will be small, often fewer than 5; therefore some form of borrowing between time and strata is required.
The statistical model and inference procedure should have the following properties.
\begin{enumerate}
    \item The smoothing model over time should penalise changes from continuous exponential growth.
        Penalising changes from continuous exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalisation has previously been shown to improve estimates of SARS-CoV-2 prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}.
    \item Limit estimates to epidemiologically and biologically plausible values.
        That is, incidence proportion and prevalence should be in the interval $[0, 1]$.
        % While incidence proportion can theoretically be greater than 1, this would imply that some individuals were infected multiple times in a day which is not biologically plausible.
    \item Have sufficient parameters to incorporate all necessary covariates to correct for non-response bias (see \cref{E-biology-data:sec:cis}).
    \item Borrow information across the different covariates.
        The borrowing should penalise towards the distribution of incidence/prevalence across strata in the same geographic region being constant.
        Standard stratified, mechanistic transmission model (such as the one introduced in \cref{E-SEIR:sec:structured-populations}) predict this, if behavioural changes are independent of strata and there is negligible impact of immunity.
    \item Be computationally efficient.
        The model will be fit to granular data and therefore needs to be run within the SRS, with its computational limitations (see\todo{ref section on SRS computational limitations}).
\end{enumerate}

@@

I link the modelled proportion of individuals that are RT-PCR positive to the data using a beta-binomial likelihood (as discussed in \cref{E-biology-data:sec:clustering}).
If on each day $t$, each strata $i$ is observed to have $y_{at}$ positives out of $n_{at}$ tests, then the likelihood is:
\begin{align}
    \prod_{a,t} \BB (y_{at} \mid n_{at}, \bar{p}_{at}, \rho)
    \label{SEIR:eq:likelihood}
\end{align}
where: $\bar{p}_{at}$ is the mean proportion of individuals in strata $a$ that are RT-PCR positive on day $t$; and $\rho$ is a nuisance parameter controlling the overdispersion of the beta-binomial distribution.
Denoting the proportion of stratum $a$ in state $p_k$ as $p_{ka}(\tau)$, then $\bar{p}_{at} = \int_{t}^{t+1} \sum_{k=1}^l p_{ka}(\tau) d\tau$.
